---
title: "Soilspec_parte1"
author: "José Ramón Cuesta"
date: "29/5/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Cargando los paquetes necesarios
```{r}
library(prospectr)
library(caret)
library(chemometrics)
```

## Preparación de los datos
```{r}
parameters<-as.matrix(datsoilspc[,1:4])
spectra<-as.matrix(datsoilspc$spc)    # The spectra is in reflectance
spc_abs<- apply(spectra, 2, log1R)    # spectra converted to Absorbance
```

## Visualización de los espectros
```{r}
#en reflectancia
matplot(wavelengths, t(spectra), type = "l", 
        xlab = "wavelength", ylab = "Reflectance")
# en absorbancia
matplot(wavelengths, t(spc_abs), type = "l", 
        xlab = "wavelength", ylab = "log 1/R")
```

## Aplicando los tratamientos matemáticos
Estos son unos de los tratamientos matemáticos más comunes que utilizo:
### SNV Only
```{r}
spectra_snv<-scale(t(spc_abs),center=TRUE,scale=TRUE)
spectra_snv<- t(spectra_snv)
colnames(spectra_snv)<- wavelengths
matplot(as.numeric(colnames(spectra_snv)), t(spectra_snv), type = "l",         xlab = "wavelength", ylab = "Log 1/R")
```
### SNV and Detrend
```{r}
spectra_snvdt<- detrend(spc_abs, as.numeric(colnames(spectra)))
matplot(as.numeric(colnames(spectra_snvdt)), t(spectra_snvdt), 
        type = "l", col = "blue", xlab = "wavelength", 
        ylab = "Log 1/R")
```
### SNV+DT+1D SPECTRA
```{r}
spectra_1d<-gapDer(spectra_snvdt,m=1,w=21,s=4)
matplot(as.numeric(colnames(spectra_1d)),t(spectra_1d) ,type="l",
        col = "blue", xlab = "wavelength", ylab = "Log 1/R")
```
### SNV + 2D SPECTRA
```{r}
spectra_2d<-gapDer(spectra_snv,m=2,w=21,s=4)
matplot(as.numeric(colnames(spectra_2d)), t(spectra_2d) ,type="l",
        col = "blue", xlab = "wavelength", ylab = "Log 1/R")
```
## CORRELATION OF SPECTRA & PARAMETERS
### Math treatment spectra correlation with Clay
Spectra is the same for all the parameters, so we change de column number of the parameter.
```{r}
cor_specsnv_clay<-cor(parameters[,1],spectra_snv[,1:ncol(spectra_snv)])
cor_specsnvdt_clay<-cor(parameters[,1],spectra_snvdt[,1:ncol(spectra_snvdt)])
cor_spec1d_clay<-cor(parameters[,1],spectra_1d[,1:ncol(spectra_1d)])
cor_spec2d_clay<-cor(parameters[,1],spectra_2d[,1:ncol(spectra_2d)])
matplot(as.numeric(colnames(spectra_snv)),t(cor_specsnv_clay),
        type = "l",xlab="nm",col = "red", ylab="Correlation",
        ylim = c(-1, 1),xlim = c(350,2500),
        main = "Mathtreatments Correlations with Clay")
par(new = TRUE)
matplot(as.numeric(colnames(spectra_snvdt)),t(cor_specsnvdt_clay),
        type = "l",xlab=" ",col = "brown", ylab=" ",
        main = " ", ylim = c(-1, 1), xlim = c(350,2500))
par(new = TRUE)
matplot(as.numeric(colnames(spectra_1d)),t(cor_spec1d_clay),
        type = "l",xlab=" ",col = "green", ylab=" ",
        main = " ", ylim = c(-1, 1),xlim = c(350,2500))
par(new = TRUE)
matplot(as.numeric(colnames(spectra_2d)),t(cor_spec2d_clay),
        type = "l",xlab=" ",col = "orange", ylab=" ",
        main = " ", ylim = c(-1, 1),xlim = c(350,2500))
```

## LOOKING TO THE "Y" MATRIX
### HISTOGRAMS
```{r}
hist(parameters[,1], main = "Clay")
hist(parameters[,2], main = "Sand")
hist(parameters[,3], main = "Silt")
hist(parameters[,4], main = "Total Carbon")
```

### CORRELATION MATRIX
```{r}
library(corrplot)
cor<-cor(parameters)
corrplot(cor, method = "circle")
```

## Principal Components Analysis with CARET
```{r}
spc_snvdt_pca<-preProcess(spectra_snvdt,
                       method = c("center", "scale","pca"),
                       thresh = 0.95)
PC_scores<-predict(spc_snvdt_pca,spectra_snvdt)
plot(PC_scores[,1],PC_scores[,2],col="blue",
     xlim=c(min(PC_scores[,1]),max(PC_scores[,1])),
     ylim = c(min(PC_scores[,2]),max(PC_scores[,2])),
     xlab = "PC1",ylab = "PC2")
```

###standard deviation & mean spectra
```{r}
matplot(wavelengths,spc_snvdt_pca$std, type = "l", col = "blue")
par(new=TRUE)
matplot(wavelengths,spc_snvdt_pca$mean, type = "l", col = "red")
```
### Loadings
```{r}
matplot(wavelengths,spc_snvdt_pca$rotation[,5], type = "l", col = "blue")
```
### Plotting projections
```{r}
matplot(seq(1,5,by=1),t(PC_scores),pch=1, col = "blue",
        xlab = "PCs", ylab = "Projections")
```

